Demographic Factors and Climate Action: Understanding How Age and Education Shape Sustainable Choices

### 2024-12-14 paper/climate_change.qmd
nice rending in data section

The remainder of this paper is structured as follows: @sec-data describes the data and methodology used for analysis, @sec-model outlines the statistical model used to assess the effects of age and education on behavior adoption, @sec-results presents the key findings, and @sec-discussion explores the implications of these results, addresses limitations, and suggests directions for future research. Finally, @sec-conclusion summarizes the main takeaways and offers recommendations for enhancing climate action engagement. Additional methodological details and the data cleaning process are provided in @sec-appendix.

# Data {#sec-data}
The primary data source for this project is the Climate Perception Study dataset, provided by the City of Toronto via the Open Data Toronto platform [@climateperceptionstudy]. This dataset offers insights into public perceptions of climate change in Toronto in the years 2018 and 2021. This dataset was accessed from `Open Data Toronto` [@opendatatoronto] on 6 November 2024.

@@

In the download script, `httr`[@httr] retrieved files from `Open Data Toronto` [@opendatatoronto], while`readxl` [@readxl] read Excel files, and `openxlsx` [@openxlsx] created structured Excel workbooks. The `arrow` package [@arrow] optimized data storage and was used in the simulation and cleaning scripts to save datasets as Parquet files.

@@

## Features
### Age
#### Individual 2018
fig-one presents the age distribution of survey respondents in 2018. The histogram shows the frequency of respondents across age groups, with age intervals on the x-axis and the number of respondents on the y-axis. The distribution is fairly balanced, with peaks around the 30–40 and 50–60 age ranges, indicating that these groups were the most represented in the survey.

@@

```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-one
#| fig-cap: "Distribution of age given the 2018 survey respondents. The mean and median of the sample is indicated with dashed and dotted lines respectively."

@@

```
\newpage

#### Comparison 2018 to 2021
fig-two compares the age distributions between 2018 and 2021. The main difference is an increase in respondents aged 18-23 in 2021, with other age groups remaining relatively stable. This shift may reflect changing demographic patterns or differences in survey methodologies within the three years.

```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-two
#| fig-cap: "This table highlights the age distribution of survey respondents in 2018 and 2021."

@@

# Print the cleaned table (using kable for styling)
age_summary_combined %>%
  kbl() %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

@@

### Education
#### Individual 2018
fig-three shows the distribution of respondents by their highest level of education attained in 2018. The most common education levels are "Completed undergraduate degree" and "Postgraduate/professional school," indicating a well-educated respondent population. In contrast, fewer respondents reported having attended or completed community college, vocational, or trade school, suggesting lower representation in these categories. A small portion of respondents chose not to disclose their education level. Overall, the distribution reflects a skew toward higher levels of educational attainment.
```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-three
#| fig-cap: "The histogram shows the distribution of education levels of 2018 survey respondents."

knitr::include_graphics(here("data/03-figures_data", "educ_individual_plot.png"))

#### Comparison 2018 to 2021
fig-four compares the education levels of respondents in 2018 and 2021. The two surveys used different education categories, making direct comparisons challenging. In 2018, 9% of respondents had a high school diploma or less, while in 2021, this figure rose to 38%, due to the addition of the "less than high school" and "high school or equivalent" categories. This shift indicates that the 2021 survey better represented individuals with lower educational attainment.

@@

```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-four
#| fig-cap: "This table highlights the education distribution of survey respondents in 2018 and 2021."

# Load necessary libraries
library(tinytable)

@@

# Convert the data frame to tinytable
education_summary_combined_tiny <- tt(education_summary_combined)

# Print the tinytable
education_summary_combined_tiny

@@

### Extent Feel Informed
#### Individual 2018
fig-five depicts the distribution of self-reported climate change knowledge among survey respondents in 2018. The majority of participants identified as "Very informed," making this the most prevalent category by a significant margin. A moderate number of respondents reported being "Not very informed," while a smaller yet noticeable group described themselves as "Extremely informed." In contrast, the categories "Not at all informed" and "Not very informed" have the lowest frequencies, with "Not at all informed" being particularly rare. This distribution suggests that most respondents possess a relatively high level of awareness about climate change, with very few indicating a lack of knowledge on the topic.
```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-five
#| fig-cap: "The histogram shows the distribution of self-reported climate change knowledge among 2018 survey respondents."

@@

```
#### Comparison 2018 to 2021
fig-six illustrates the self-reported level of extent inviduals feel informed about climate change and climate action for 2018 and 2021. In 2018, the large marjority of individuals say they are very informed with 60 percent and the follwoign category is not very informed with 27 percent. In 2021, The majority of inividuals also say they are very informed but to a lesser degree with 48 percent and this is closely followed by only a little informed with 37 percent. There is a trend that invidiusls in 2018 reported greater levels of knowldege about climate change whereas 2021 illstrates a more even split between very informed and only little informed, portraying that fewer individuals have adequate knowledge about climate change.

```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-six
#| fig-cap: "This table highlights the extent feel informed distribution of survey respondents in 2018 and 2021."

@@

# Print the table (using kable for styling)
informed_summary_combined %>%
  kbl() %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

@@

### Best Method of Delivery
#### Individual 2018
fig-seven, shows 2018 survey results for the preferred method for receiving climate change information. Advertising campaigns emerged as the most popular choice, selected by approximately 55 percent of respondents, followed closely by newsletters with 52 percent and the Toronto.ca website with 46 percent. On the other hand, social media platforms such as Facebook, Instagram, and Twitter were less popular overall, with Twitter 12 percent and Instagram 17 percent garnering the fewer votes. Notably, only 5 percent of respondents reported being uninterested in receiving information about climate action. These findings suggest that more traditional communication methods, such as advertising and newsletters, resonated better with respondents compared to social media.

>>>>

The remainder of this paper is structured as follows: The `data section` describes the data and methodology used for analysis, the `model section` outlines the statistical model used to assess the effects of age and education on behavior adoption, the `results section` presents the key findings, and teh `discussion section` explores the implications of these results, addresses limitations, and suggests directions for future research. Finally, the `conclusion section` summarizes the main takeaways and offers recommendations for enhancing climate action engagement. Additional methodological details and the data cleaning process are provided in the `appendix section`.

# Data 
The primary data source for this project is the Climate Perception Study dataset, provided by the City of Toronto via the Open Data Toronto platform [@climateperceptionstudy]. This dataset offers insights into public perceptions of climate change in Toronto in the years 2018 and 2021. This dataset was accessed from `Open Data Toronto` [@opendatatoronto] on 6 November 2024.

@@

In the download script, `httr` [@httr] retrieved files from `Open Data Toronto` [@opendatatoronto], while `readxl` [@readxl] read Excel files, and `openxlsx` [@openxlsx] created structured Excel workbooks. The `arrow` package [@arrow] optimized data storage and was used in the simulation and cleaning scripts to save datasets as Parquet files.

@@

## Features
### Age
#### Individual 2018
Figure 1 presents the age distribution of survey respondents in 2018. The histogram shows the frequency of respondents across age groups, with age intervals on the x-axis and the number of respondents on the y-axis. The distribution is fairly balanced, with peaks around the 30–40 and 50–60 age ranges, indicating that these groups were the most represented in the survey.

@@

```{r}
#| warning: false
#| echo: false
#| message: false
#| fig-cap: "Distribution of age given the 2018 survey respondents. The mean and median of the sample is indicated with dashed and dotted lines respectively."

@@

#### Comparison 2018 to 2021
Table 1 compares the age distributions between 2018 and 2021. The main difference is an increase in respondents aged 18-23 in 2021, with other age groups remaining relatively stable. This shift may reflect changing demographic patterns or differences in survey methodologies within the three years.

```{r}
#| warning: false
#| echo: false
#| message: false
#| fig-pos: "H"  # Force figure placement here

@@

# Render the table with a manually added caption
age_summary_combined %>%
  kbl(caption = "This table highlights the age distribution of survey respondents in 2018 and 2021.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

@@

### Education
#### Individual 2018
@fig-educ shows the distribution of respondents by their highest level of education attained in 2018. The most common education levels are "Completed undergraduate degree" and "Postgraduate/professional school," indicating a well-educated respondent population. In contrast, fewer respondents reported having attended or completed community college, vocational, or trade school, suggesting lower representation in these categories. A small portion of respondents chose not to disclose their education level. Overall, the distribution reflects a skew toward higher levels of educational attainment.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-educ
#| fig-cap: "The histogram shows the distribution of education levels of 2018 survey respondents."

knitr::include_graphics(here("data/03-figures_data", "educ_individual_plot.png"))
```

#### Comparison 2018 to 2021
Table 2 compares the education levels of respondents in 2018 and 2021. The two surveys used different education categories, making direct comparisons challenging. In 2018, 9% of respondents had a high school diploma or less, while in 2021, this figure rose to 38%, due to the addition of the "less than high school" and "high school or equivalent" categories. This shift indicates that the 2021 survey better represented individuals with lower educational attainment.

@@

```{r}
#| warning: false
#| echo: false
#| message: false
#| fig-cap: "Table 2: This table highlights the education distribution of survey respondents in 2018 and 2021."

# Load necessary libraries
library(tinytable)
library(kableExtra)
library(here)

@@

# Render the table with adjusted font size and column width
education_summary_combined %>%
  kable() %>%
  kable_styling(full_width = TRUE, position = "center") %>%
  column_spec(1, width = "10em") %>%
  column_spec(3, width = "10em") %>%
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 8) %>%
  # Add resizebox to scale the table down to fit within the page
  row_spec(0, bold = TRUE) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)

# Print the tinytable
#education_summary_combined

@@

### Extent Feel Informed
#### Individual 2018
Figure 3 depicts the distribution of self-reported climate change knowledge among survey respondents in 2018. The majority of participants identified as "Very informed," making this the most prevalent category by a significant margin. A moderate number of respondents reported being "Not very informed," while a smaller yet noticeable group described themselves as "Extremely informed." In contrast, the categories "Not at all informed" and "Not very informed" have the lowest frequencies, with "Not at all informed" being particularly rare. This distribution suggests that most respondents possess a relatively high level of awareness about climate change, with very few indicating a lack of knowledge on the topic.
```{r}
#| warning: false
#| echo: false
#| message: false

@@

```
#### Comparison 2018 to 2021
Table 3 illustrates the self-reported level of extent individuals feel informed about climate change and climate action for 2018 and 2021. In 2018, the large majority of individuals say they are very informed with 60 percent and the following category is not very informed with 27 percent. In 2021, The majority of individuals also say they are very informed but to a lesser degree with 48 percent and this is closely followed by only a little informed with 37 percent. There is a trend that individuals in 2018 reported greater levels of knowledge about climate change whereas 2021 illustrates a more even split between very informed and only little informed, portraying that fewer individuals have adequate knowledge about climate change.

```{r}
#| warning: false
#| echo: false
#| message: false

@@

# Print the table (using kable for styling)
informed_summary_combined %>%
  kbl(caption = "This table highlights the extent feel informed distribution of survey respondents in 2018 and 2021") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

@@

### Best Method of Delivery
#### Individual 2018
Figure 4 shows 2018 survey results for the preferred method for receiving climate change information. Advertising campaigns emerged as the most popular choice, selected by approximately 55 percent of respondents, followed closely by newsletters with 52 percent and the Toronto.ca website with 46 percent. On the other hand, social media platforms such as Facebook, Instagram, and Twitter were less popular overall, with Twitter 12 percent and Instagram 17 percent garnering the fewer votes. Notably, only 5 percent of respondents reported being uninterested in receiving information about climate action. These findings suggest that more traditional communication methods, such as advertising and newsletters, resonated better with respondents compared to social media.
